Speculation in foreign exchange: noise or information

### 2015-07-13 Knitr/NewPaper.Rnw
Completed writing NewPaper.Rnw There is still a problem with the appendix.

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing the the strength of speculative opinion or the weight of speculative positions does not improve forecasts of future returns. This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.

@@

% does this following paragraph come later?  I think so....Once the three
% main categories are done. 
In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.  They do not usually distinguish between psychological and institutional reasons for the lack of attention to fundamental information. Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value. This may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninformed speculation provides a rational and an opportunity for the trading of financial securities. If all economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, the informed can take advantage of their information.  This creates a justification for acquiring information. Speculative noise-traders can therefore create a gap between the price and the fundamental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundamental value is thereby essential to encourage the process of price discovery. 
%
 \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

  Speculation is also an important part of the argument for why prices should tend towards a level that consistent with \emph{the rational expectations equilibrium} (RER). This is the level that is consistent with rational expectations.  In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent. These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the means by which the components of the model tend towards equilibrium values.  
% See \citet gc{Lyons2006Microstructure} and \citep{Evans2002Order} for examples of the micrstrucure approach in foreign exchange.

However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inappropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements.

The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.  Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than actual underlying, fundamental value \citet[p.101]{Keynes1936}

it has also been suggested that speculative techniques can change overtime.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation. 
%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
%this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
\subsection{Empirical work}
Empirical assessment of speculation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish those that may be noise-traders from those assumed to be informed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  For example speculation can be identified with momentum trading. It is also possible to assume that certain investors, such as investment banks and hedge funds, are informed while other trade without information. The latter group may include day-traders or private individuals.

%The literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.  These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.

\subsection{Why this is important}
The Keynesian view of speculation is that it is an addition to market noise that makes it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. This description provides an explanation for Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

  The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).  There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.  Regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they reduce risk and encourage more long-term attitudes towards investment.  However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.

@@

%In addition, so-called \emph{Volcker Rule} would prevent US investment banks trading on their own-account.  This, it has been argued, will reduce liquidity in the bond market by causing investment banks to hold lower levels of bond inventory on their books.  This may increase the amount of \emph{price slippage} when final investor adjust their positions. In addition,

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the  foreign exchange market. This will be use to gain a fuller understanding of speculative behaviour.   The experiment will involve three steps:  assess the price implication of the noise-trader model;  measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.  
% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citet gc{FrootChartists} for a survey.  %this may go in the explanation for why fx is used.  Use this elsewhere. Add more?

The \citet{Delong1990noise} noise-trader model used in this study. This is a derivative of the \citet{Samuelson1958} overlapping generations model  with two-period agents making a portfolio choice.  There are two assets:  a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; the risky asset $(u)$ also pays the same dividend but is not in elastic supply.  There are two types of investor or trader: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

@@

the final price equation becomes.

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}
\label{eq:NP2}
\end{equation}

@@

The Friedman-effect and the Price-pressure are the effects that are studied here.  Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversals.  The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.  Create-space effect would suggest that when the weight of noise-traders is high, there is increased risk.  %this can be assessed.

The main findings are that noise-traders will not necessarily disappear as Friedman had suggested \citet{FriedmanPositive}.  While noise-traders will tend to get less utility than sophisticated traders, this does not necessarily drive them from the market.  The noise-traders expected to get higher utility due to their misperception.  The reduction in utility comes mainly from the fact that they are not compensated for taking risk that they are not aware of.  This is consistent with the type of over-confidence that has been found by \citet{odean1998volume}in day-traders.

\subsection{The implications of the noise-trader model}
There are two parameters from the noise-trader price equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  the aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.

@@

% this needs to be completed.

@@

% The table may be pushed onwards.  Use t (i think).

Table \ref{tabref:DS1} in Appendix \ref{apref:1} shows descriptive statistics for the exchange rate data used in the study.  There are daily data for the period between January 1996 and September 2002.  This is the period for which risk-reversal data are available.  The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.  From the raw data continuously compounded returns are calculated and multiplied by 100.  The first thing of note is that the mean return does not appear to be statistically different from zero.  This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk. 
\begin{sidewaystable}
\begin{threeparttable}
\caption{Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
\begin{tabular}{l | r r  r r r r r r r r}
  \hline
 Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
 \hline
Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
 Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
 Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
 Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
      \hline
 \label{tabref:DS2}
\end{tabular}
\begin{tablenotes}
\small
\item Daily risk reversal skew.  Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).  A figure above zero means that calls are more expensive (in volatility and price terms) than puts.  A figure below zero means that puts are more expensive. The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.   These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{sidewaystable}

@@

%The operation of derivative exchanges is a competitive business that has evolved over the course of the data collection period.  The two original exchanges in Chicago, the Chicago Mercantile Exchange (CME) and the  Chicago Board of Trade (CBOT) went public in 2000 and 2005 respectively and merged in 2007 to become the CEM Group.  In 2008 the New York Mercantile Exchange (NYMEX) was added,  Most of the successful currency derivative contracts were traded on the International Monetary Market (IMM) which was a sub-division of the original CME.  Many exchange rates and cross rate products were launched by these and other US exchanges, but the five contract which maintained a presence and liquidity through the whole period were the Deutschmark and then the Euro, the Swiss Franc, the British Pound, the Japanese Yen and the Canadian Dollar.  There is a reasonably long series for the Australian Dollar and for the Mexican Peso, but nether of these is complete.  The US dollar index has been active since it was launched in May 1992 but it is not trade weighted or really representative of the US dollar.

The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.  This date is shortly before this part of the research was started. The Euro data runs  from March 3 1998 to December 31 2008.  There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.  The absence of these data seems to be a result of problems with the CFTC database.  The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.

@@

Looking down the seventh column (the 4-day CARW) shows the average cumulative return for the period from four days before the extreme event to the four days after the extreme for extreme sentiment highs and lows of each currency pair.   The extreme highs are (in this case) those points when the risk-reversal skew is at or above the 95th percentile.  It should not be surprising to see that the extreme highs are associated with positive returns, which appear in many cases to be significantly different from zero (where significance is identified as more than 95\% of 1000 random bootstrap samples taken from the events being above zero).  
It is clear that, as speculators become more extreme in their opinion, prices are driven in the direction of sentiment.  It shows that the spot price is following the option risk-reversal skew and, as Figure \ref{fig:ES1} reveals, it would be even more pronounced if the event window just ran from the starting point to the event.  The findings here are consistent with the evidence that orders and sentiment drive prices \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.

@@

However, it does indicate that the evolution of exchange rates are not a random walk.  There is evidence here that prices gradually respond to information, as would be the case if decision-making is following a \emph{conservatism} heuristic or if there were institutional features.  The positive (or negative) expectations that are evident in the risk reversal skew do not seem to represent a deviation or a mis-pricing that will be reversed.  Rather they seem to be part of the price-discovery process.

@@

There is strong evidence that speculation affects prices and that this effect is gradual rather than instantaneous.  This consistent with a market inefficiency that would reward momentum-based trading.  It is no wonder therefore that \citet{TaylorTechnical}, reporting a survey conducted by the Bank of England, found that of the Chief foreign exchange dealers contacted in London in November 1988, 90\% of them attached some weight to \emph{technical analysis}, a method that is fundamentally based on finding trends in asset prices.  The survey found that the importance of technical analysis became greater at shorter horizons.  This has not disappeared.  The importance of momentum trading and technical analysis has not waned with time.  Neely and Weller  show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases \citet{NeelyWeller}. It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected in the purest form of the EMH.

The notion that speculation is at least partially informed is supported by the fact that there is little evidence from the event studies conducted here that speculation causes overshooting or bubbles.  When speculation is extreme, whether measured by intensity of belief or weight of speculative positions, future returns are as likely to be positive as negative.  There is no informational content as there would be if speculation was pure noise.  In the simulation of the De Long model, extremes are more likely to be followed by reversals.  In the foreign exchange market, that is not the case.  There are some large price crashes in the dataset, as shown by the descriptive statistics of the returns. For example  in Table \ref{tabref:DS1} it is seen that on one day there is a 10.65\% appreciation of the Japanese yen and Table \ref{tabref:DS2} shows a week when the Canadian dollar falls by 10\%.  However, the measures of speculative sentiment or the weight of speculation are not particularly high in either of these cases.  Indeed, an analysis of the top 1\% of positive and negative foreign exchange returns in the dataset reveals that sentiment and positions for this group were very close to the average of the whole set.  This would be consistent with a model where extreme price movement was a function of new information rather than the structure of existing positions.    
%expand.  Could put data in appendix.

The evidence here, that speculation is associated with serial correlation in the returns but cannot be used in a systematic way to determine reversals, has important implications for speculative activity.  It means that the largest returns will be to those that identify the trend at the earliest opportunity.  This means that speculators must be prepared to act as swiftly as possible, probably when there is a high level of uncertainty and a lack of information and where any scrap of additional news is highly valued.   The risk that actions by other speculators, most notable by their effect on price, will be construed to contain information exacerbates the likelihood of feedback effects.  The need to be able to take action before others has reached the most extreme form in the activity of \emph{high frequency trading} (HFT), where decision times are measured in nanoseconds and there is advantage in gaining spacial proximity to exchanges.  HFT is all about getting  a larger share of the inefficiency in pricing information by trading or investing closer to where the information is released.

@@

\begin{sidewaystable}
\begin{threeparttable}
\caption{Descriptive Statistics for Weekly Exchange Rate Returns: 30 September 1998 to 31 December 2008}
\begin{tabular}{{4cm} | r r  r r r }
  \hline
 Statistic &  EUR & JPY & GBP & CHF & CAD  \\
  \hline
Number & 522 &  849 & 849 & 847 & 849 \\
Mean & 0.0351 & 0.0337 & -0.0202 & 0.0233 & 0.0018\\
Median & 0.0650 & -0.0800 & 0.0000 & 0.0000 & -0.0200\\
Standard deviation  & 1.4063 & 1.6355 & 1.2755 & 1.5418 & 1.0558\\
Max & 8.7900 & 11.6700 & 5.7200 & 8.3400 & 10.0100\\
Min & -4.2400 & -15.0400 & -6.4100 & -5.7100 & -5.0900\\
Skew  & 0.27* & -0.06 & -0.34* & 0.31* & 0.78*\\
Kurtosis & 2.76*& 12.58* & 2.07 * & 1.82 & 11.59*\\
\hline
 \label{tabref:DS3}
\end{tabular}
\begin{tablenotes}
\small
\item Continuously compounded returns multiplied by 100 for US dollar per unit of Euro (EUR), US dollar per unit of Japanese Yen (JPY), US dollar per unit of UK Sterling (GBP), US dollar per unit of Swiss Franc (CHF) and US dollar per unit of Canadian dollar (CAD) for the period 30 September 1998 to 31 Dec 2008;     An asterisk (*) means that the estimate is significantly different from zero at the 5\% level of statistical significance using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.  These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{sidewaystable}

>>>>

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing  the strength of speculative opinion or the weight of speculative positions does not improve forecasts of future returns. This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence seems to help rather than hinder the incorporation of information into prices.

@@

In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.  They do not usually distinguish between institutional and psychological reasons for the lack of attention to fundamental information. Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value. This may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninformed speculation provides a rational and an opportunity for the trading of financial securities. If all economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, the informed can take advantage of their information.  This creates a justification for acquiring information. Speculative noise-traders can therefore create a gap between the price and the fundamental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundamental value is thereby essential to encourage the process of price discovery. 
%
 \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value. This begs the question: are speculators trading on noise or are they informed?

  Speculation is also an important part of the argument for why prices should tend towards a level that consistent with \emph{the rational expectations equilibrium} (RER). This is the level that is consistent with rational expectations.  In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent. These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the mechanisem through which equilibrium is achieved.

However, though equating noise and noise-trading as a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, it may also be an over-simplification that is inappropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speuclative activity.

The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.  Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than actual underlying, fundamental value \citet[p.101]{Keynes1936}.

\subsection{Empirical work}
Empirical assessment of speculation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish those that may be noise-traders from those assumed to be informed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  For example speculation can be identified with momentum trading. It is also possible to assume that certain investors, such as investment banks and hedge funds, are informed while other trade without information. The latter group may include day-traders or private individuals.  
It has also been suggested that speculative techniques can change over time.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.

%Noise stragegies may be successful for a while.  Momentum trades tend to be successful for certain ties.  Also discussion the literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.  These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.

\subsection{Why this is important}
The Keynesian view of speculation is that it is an addition to market noise that makes it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. As such, it exacerbates the tendency towards putting too much emphasis on the immediate and works against longer-term and strategic thinking.  This description provides an explanation for Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).  There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.  For example, regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they reduce risk and encourage more long-term attitudes towards investment.  However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.

@@

%volcker rule?

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the foreign exchange market. This will be used to gain a fuller understanding of speculative behaviour.   The analysis will involve three steps:  assessement of the price implication of the noise-trader model;  measurement of speculative activity in the foreign exchange market; finally, assessement of whether extreme speculative market activity is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.

The \citet{Delong1990noise} noise-trader model used in this study. This is a derivative of the \citet{Samuelson1958} overlapping generations model with two-period agents making a portfolio choice between two assets: a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; a risky asset $(u)$ also pays the same dividend but is not in elastic supply.  There are two types of investor: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

@@

the final price equation becomes:

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
\label{eq:NP2}
\end{equation}

@@

The Friedman-effect and the Price-pressure are the effects that are studied here.  Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversal.  The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.  Create-space effect would suggest that when the weight of noise-traders is high, there is increased risk.  %this can be assessed.

\subsection{The implications of the noise-trader model}
There are two parameters from the noise-trader price equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  The aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.

@@

@@

Table \ref{tabref:DS1} in Appendix \ref{apref:1} shows descriptive statistics for the exchange rate data used in the study.  There are daily data for the period between January 1996 and September 2002.  This is the period for which risk-reversal data are available.  The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.  From the raw data continuously compounded returns are calculated and multiplied by 100.  The first thing of note is that the mean return does not appear to be statistically different from zero.  This is what would be expected if, as has been identified in many studies, exchange rates follow something close to a random walk. 
\begin{sidewaystable}
\begin{threeparttable}
\caption{Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
\begin{tabular}{l | r r  r r r r r r r r}
  \hline
 Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
 \hline
Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
 Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
 Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
 Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
      \hline
 \label{tabref:DS2}
\end{tabular}
\begin{tablenotes}
\small
\item Daily risk reversal skew.  Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).  A figure above zero means that calls are more expensive (in volatility and price terms) than puts.  A figure below zero means that puts are more expensive. The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.   These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{sidewaystable}

@@

The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.  The Euro data runs from March 3 1998 to December 31 2008.  There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.  The absence of these data seems to be a result of problems with the CFTC database.  The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.

@@

Looking down the seventh column (the 4-day CARW) shows the average cumulative return for the period from four days before the extreme event to the four days after the extreme for extreme sentiment highs and lows of each currency pair.   The extreme highs are (in this case) those points when the risk-reversal skew is at or above the 95th percentile.  It should not be surprising to see that the extreme highs are associated with positive returns, which appear in many cases to be significantly different from zero (where significance is identified as more than 95\% of 1000 random bootstrap samples taken from the events being above zero).

It is clear that, as speculators become more extreme in their opinion, prices are driven in the direction of sentiment.  It shows that the spot price is following the option risk-reversal skew and, as Figure \ref{fig:ES1} reveals, it would be even more pronounced if the event window just ran from the starting point to the event.  The findings here are consistent with the evidence that orders and sentiment drive prices \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}.

@@

However, it does indicate that the evolution of exchange rates are not a random walk.  There is evidence here that prices gradually respond to information, as would be the case if there were institutional or psychological features that caused momentum or gradual adjustment. The positive (or negative) expectations that are evident in the risk reversal skew do not seem to represent a deviation or a mis-pricing that will be reversed.  Rather they seem to be part of the price-discovery process.

@@

There is strong evidence that speculation affects prices and that this effect is gradual rather than instantaneous.  This consistent with a market inefficiency that would reward momentum-based trading.  It is no wonder therefore that \citet{TaylorTechnical}, reporting a survey conducted by the Bank of England, found that of the Chief foreign exchange dealers contacted in London in November 1988, 90\% of them attached some weight to \emph{technical analysis}, a method that is fundamentally based on finding trends in asset prices.  The survey found that the importance of technical analysis became greater at shorter horizons.  This has not disappeared.  The importance of momentum trading and technical analysis has not waned with time.  The more up-to-date CFTC data show the same pattern.  In addition,  \citet{NeelyWeller} show that gains from technical trading rules still generate excess returns and attribute this to behavioural biases. It seems that speculation forms part of the process of price discovery, but this process is not instantaneous as would be expected in the purest form of the EMH.

The notion that speculation is at least partially informed is supported by the fact that there is little evidence from the event studies conducted here that speculation causes overshooting or bubbles.  When speculation is extreme, whether measured by intensity of belief or weight of speculative positions, future returns are as likely to be positive as negative.  There is no informational content as there would be if speculation was pure noise.  In the simulation of the De Long model, extremes are more likely to be followed by reversals.  In the foreign exchange market, that is not the case.  There are some large price crashes in the dataset, as shown by the descriptive statistics of the returns. For example  in Table \ref{tabref:DS1} it is seen that on one day there is a 10.65\% appreciation of the Japanese yen and Table \ref{tabref:DS2} shows a week when the Canadian dollar falls by 10\%.  However, the measures of speculative sentiment or the weight of speculation are not particularly high in either of these cases.  Indeed, an analysis of the top 1\% of positive and negative foreign exchange returns in the dataset reveals that sentiment and positions for this group were very close to the average of the whole set.  This would be consistent with a model where extreme price movement was a function of new information rather than the structure of existing positions.

@@

\begin{threeparttable}
\caption{Descriptive Statistics for Weekly Exchange Rate Returns: 30 September 1998 to 31 December 2008}
\begin{tabular}{{4cm} | r r  r r r }
  \hline
 Statistic &  EUR & JPY & GBP & CHF & CAD  \\
  \hline
Number & 522 &  849 & 849 & 847 & 849 \\
Mean & 0.0351 & 0.0337 & -0.0202 & 0.0233 & 0.0018\\
Median & 0.0650 & -0.0800 & 0.0000 & 0.0000 & -0.0200\\
Standard deviation  & 1.4063 & 1.6355 & 1.2755 & 1.5418 & 1.0558\\
Max & 8.7900 & 11.6700 & 5.7200 & 8.3400 & 10.0100\\
Min & -4.2400 & -15.0400 & -6.4100 & -5.7100 & -5.0900\\
Skew  & 0.27* & -0.06 & -0.34* & 0.31* & 0.78*\\
Kurtosis & 2.76*& 12.58* & 2.07 * & 1.82 & 11.59*\\
\hline
 \label{tabref:DS3}
\end{tabular}
\begin{tablenotes}
\small
\item Continuously compounded returns multiplied by 100 for US dollar per unit of Euro (EUR), US dollar per unit of Japanese Yen (JPY), US dollar per unit of UK Sterling (GBP), US dollar per unit of Swiss Franc (CHF) and US dollar per unit of Canadian dollar (CAD) for the period 30 September 1998 to 31 Dec 2008;     An asterisk (*) means that the estimate is significantly different from zero at the 5\% level of statistical significance using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.  These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
